Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks

### 2023-01-06 paper-1/manuscript.tex
Rev: F7 text minor

Comparing the non-target response reductions in Figure \ref{fig7}b, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians differing by 36\%. Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials. Indeed, the non-target response contrast remained tightly correlated with the contrast (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c. Correspondingly, $V_{TA}$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slighly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).

@@

This naturally raises the question of why, with the addition of STD, the non-target \dev{} response was reduced more than non-target \msc{} responses. We hypothesized that, due to its relatively short time constant, the direct effect of STD across trials was limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences. This would mean that STD is capable only of local suppression, reducing the standard response, but not of global suppression, leaving networks with only STD unable to differentiate between \dev{} and \msc{} target trials.

@@

To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} pulses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD. Unlike $V_{TA}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse. Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $V_{TA}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.

>>>>

Comparing the non-target response reductions in Figure \ref{fig7}b, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians differing by 36\%. Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials. Indeed, the non-target response contrast remained tightly correlated with the contrast (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c. Correspondingly, $V_{TA}$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).

@@

This naturally raises the question of why, with the addition of STD, the non-target \dev{} response was reduced more than non-target \msc{} responses. We hypothesized that, due to its relatively short time constant, the direct effect of STD across trials was limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences. This would mean that STD is capable only of local suppression, reducing the non-target response in \dev{}, but not of global suppression, leaving networks with only STD unable to differentiate between \dev{} and \msc{} target trials.

@@

To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} responses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD. Unlike $V_{TA}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse. Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $V_{TA}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
